Topological Equivariant Artist Model for Visualization Library Architecture

### 2022-04-04 paper/main.tex
fiber bundles

\section{Expressing Structure}
\label{sec:atct}
\note{why why why?}
We use concepts from algberaic topolology and category theory to formally express the structure visualization components are expected to preserve. Category theory is a mathematical langauge for describing how objects are specified, constrained, and composed \cite{wielsManagementEvolvingSpecifications1998}\; therefore in this work we propose implementable in code categorical constructions of visualization components. Fiber bundles, are a "unified, dimension-independent framework", as described by Butler\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}, that allow us to seperatly describe the topololigical properties and fields of a data source and also the mapping between topolology and data values. Sheaves on bundles are an "algebraic data structure", as described by Ghrist \cite{ghristElementaryAppliedTopology2014}, for for expressing the bookeeping involved in subsetting data and composing data that is distributed or streaming.

\subsection{Fiber Bundles}
\label{sec:atct:fiber-bundles}

>>>>

\section{Expressing Structure}
\label{sec:atct}
\note{why why why?}
We use concepts from algberaic topolology and category theory to formally express the structure visualization components are expected to preserve.  Fiber bundles and sheaves, which come from algebraic topolology, provide a uniform generalizable way to describe data and visualizations. Category theory is a method to describe how objects are specified, constrained, and composed \cite{wielsManagementEvolvingSpecifications1998}\; therefore in this work we propose categorical constructions of visualization components as specifications for implementable components.

\subsection{Fiber Bundles}
\label{sec:atct:fiber-bundles}
Fiber bundles, are a "unified, dimension-independent framework", as described by Butler\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}, that allow us to seperatly describe the topololigical properties and fields of a data source and also the mapping between topolology and data values. A fiber bundle $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and continuous surjective map $\pi: \dtotalc \rightarrow \dbasec$\cite{FiberBundle2020}.

\begin{equation}
  \label{eq:atct:fb_intro}
  \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
   \dfiberc
    \arrow[r, hook] \& 
    \dtotalc
    \arrow[d, "\pi"'] \\
     \& 
  \dbasec
     \arrow[u, "\dsectionc"', bend right, pos=.5]
  \end{tikzcd}
\end{equation}

The \textcolor{base}{base space} is a topological space \dbasec\ with points $\dbasepointc \in \dbasec$ and topology $\mathcal{T}_{\dbasec}$; $\mathcal_{T}_k$ is a set of opensets $ \dbasepointc \in \opensetc \subseteqq \dbasec$ \cite{munkresElementsAlgebraicTopology1984} that cover the base space \dbasec.  The \textcolor{fiber space} is a topological space \dfiberc\ that is the preimage of the projection function $\pi$ at a point in the base space $\dfiberc|_{\dbasepointc} = \pi^{-1}(\dbasepointc)$ and fibers of a bundle are isomorphic $\dfiberc \simeq \dfiberc_{\dbasepointc}\;\forall \dbasepointc \in \dbase$. The \textcolor{section}{sections} of a bundle $\dsection$ are maps from the base space to points in the bundles over that base space and $\textcolor{set}{\Gamma}$ denotes the set of all sections in a bundle over an openset. 
\begin{equation}
  \label{eq:atct:fb_sections}
  \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \coloneqq \big\{\gsectionc: \opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} \; \bigm{\vert} \pi(\gsectionc(\gbasepointc)) = \gbasepointc\;for\, all\; \gbasepointc \in \opensetgc \big\} 
\end{equation}

A fiber bundle is locally trivial, which means that for every point \dbasepointc\ there exists an open neighborhood $\dbasepointc \in \opensetc \subseteq \dbasec$ such that there is a homeomorphism $\pi^{-1}(\opensetc)\xrightarrow{\equivc} \opensetc \times \dfiberc$. A bundle may be globally trivial, meaning that $\dtotalc = \dbasec \times \dfiberc$, and generally can be thought of as a twisted product of total and base spaces \cite{munkresElementsAlgebraicTopology1984}.

\begin{figure}
  \label{fig:atct:fb}
  \includegraphics[width=\columnwidth]{fiberbundle.png}
  \caption{change to concrete example}
\end{figure}

Fiber bundles can be used to model data for visualization, as proposed by Butler\cite{butlerVectorBundleClassesForm1992, butlerVisualizationModelBased1989}. We encode the topological properties of the data \cite{wilkinsonGrammarGraphics2005} as the base space \dbasec. The topological properties, as described by Wilkenson, are how the data is organized - for example \dbase\ could be discrete points, a line, a 2D surface, a 3D volume, or a graph. The names and data types of the data fields can be expressed as the fiber space \dfiberc, as described by Spivak \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES}. For example, a temperature field could be expressed as a fiber $\dfiber_{temperature}=\mathbb{R}$, and a color field could be $\dfiber_{color}=\mathbb{R}^{3}$. Even though fibers on non-trivial spaces, such as vector fields, are different, we can simplify to one type because fibers are isomorphic.\note{unsure of this claim but I think this is why, and also I need to tell you why you care that they're isomorphic} The total bundle space \dtotalc\ acts as a way of expressing all data with the same topolology \dbasec\ and fields \dfiberc, for example a database schema. A specific dataset is represented as a section \dsectionc, and the return values of a section evaluated at a point \dbasepointc\ is the record $\{field_{0}: value_{0}, ..., field_n:value_n\}$ at that point.

We also use fiber bundles to represent the output of a visualization algorithm, which we term the graphic but generalizes to output on any display space, such as a screen or 3D print. 
\begin{equation}
  \label{eq:atct:fb_graphic}
  \gfiberc \hookrightarrow \gtotalc \xrightarrow{\pi} \gbasec
\end{equation}
The base space \gbasec\ is a parameterization of the display area, for example the inked bounding box in cairo \cite{CairographicsOrg}. The fiber space \gfiberc\ is an abstraction of the renderer fields, for example $\{x,\,y,\,r,\,g,\,b,\,a\}$. The sections of the graphic 
\begin{equation}
  \label{eq:atct:fb_graphic_section}
  $\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \coloneqq \big\{\gsectionc: \opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} \; \bigm{\vert} \pi(\gsectionc(\gbasepointc)) = \gbasepointc\;for\, all\; \gbasepointc \in \opensetgc \big\}$
\end{equation}
are functions that generate different graphics, for example markers for a scatter plot, bars for a bar chart, cells of a heatmap, or pixels of an image.

\subsubsection{Input and Output Types}

\subsection{Sheaves}
Sheaves on bundles are an "algebraic data structure", as described by Ghrist \cite{ghristElementaryAppliedTopology2014}, for expressing the bookkeeping involved in subsetting data and composing data that is distributed or streaming.

\subsection{Transforms on Sheaves}

\subsection{Transforms Between Sheaves}
